---
title: All World Cup Squads
author: Michael Pawlus
date: '2018-06-05'
slug: all-world-cup-squads
categories: []
tags: []
---

# World Cup 2018 is Finally Here

## A Quick Analysis of all Teams

It has been a long, long 1,423 days since the conclusion of the last World Cup. Luckily, this long wait is almost over which means it is time to analyze everything and anything World Cup related.

In this post, let's start with getting all the squads.

We will gather all the team data from the Wikipedia page.

### First, load libraries:

```{r results='hide', echo = FALSE, warning=FALSE}
# load libraries
library(tidyverse) # for data manipulation
library(rvest) # for web scraping
library(lubridate) # to work with the birthdates
```

### After, get the url for the Wikipedia page:

```{r}
url <- "https://en.wikipedia.org/wiki/2018_FIFA_World_Cup_squads"
```

There are 32 teams and each one is in a table and each table has a sequential number in the XPath for that table.

This is also true for the team names but they are in headers which also contain the same set of sequential numbers.

Since we know this, we can first create a vector of sequential numbers from 1 to 32:

```{r}
table_numbers <- c(1:32)
```

### Next, write a function to get all the player data from each table:

```{r}
# function to get table for each team
get_teams <- function(tbl_nums) {
  
  # get country names
  team_name <-   url %>%
    read_html() %>%
    html_node(xpath = paste0('//*[@id="mw-content-text"]/div/h3[',tbl_nums,']')) %>%
    html_text
  
  # get squad info
  url %>%
    read_html() %>%
    html_node(xpath = paste0('//*[@id="mw-content-text"]/div/table[',tbl_nums,']')) %>%
    html_table(fill = TRUE) %>%
    as.tibble() %>%
    mutate(team_name = team_name) # add country names
}
```

### Then, use `map` to iterate over vector of numbers and pull player data for each team:

```{r}
teams <- map_df(table_numbers, get_teams)
```

### Data Clean-Up

We will be stacking all the data for 2018 squads on top of all the historical team data for analysis later. As a result we need to make some changes to our extracted data so we can `bind_rows` with the historical data.

```{r}
# clean up the data
teams <- teams %>%
  mutate(`No.` = as.character(`No.`), ClubCountry = NA, Year = 2018) %>%
  rename(No = `No.`, Pos = `Pos.`, `DOB/Age` = `Date of birth (age)`, Country = team_name) %>%
  select(No,Pos,Player,`DOB/Age`,Caps,Club,Country,ClubCountry,Year)
```

### Get Historical Teams

Thanks to [this repo](https://github.com/sanand0/fifadata) by S. Anand we can get all historical teams:

```{r}
## read in historical data
historic_data <- read_csv("https://raw.githubusercontent.com/sanand0/fifadata/master/squads.csv")
```

### A Little More Clean-Up

We have to change the data type for the `Caps` column to match

```{r}
# change data type for Caps to match
historic_data <- historic_data %>%
  mutate(Caps = as.integer(Caps))
```

### Stack Everything

Finally, we can stack the data and we have all teams for 2018 as well as all historical teams in one table

```{r}
# stack 2018 data and historical data
all_squads <- bind_rows(historic_data,teams)
```

