---
title: All World Cup Squads
author: Michael Pawlus
date: '2018-06-05'
slug: all-world-cup-squads
categories: []
tags: []
---

# Getting All World Cup 2018 Squads

It has been a long, long 1,423 days since the conclusion of the last World Cup. Luckily, this long wait is almost over which means it is time to analyze everything and anything World Cup related.

In this post, let's start with getting all the squads.

We will gather all the team data from the Wikipedia page.

First, get the url for that page:

```{r}
url <- "https://en.wikipedia.org/wiki/2018_FIFA_World_Cup_squads"
```

There are 32 teams and each one is in a table and each table has a sequential number in the XPath for that table.

Since we know this, we can first create a vector of sequential numbers from 1 to 32

```{r}
table_numbers <- c(1:32)
```

Next, we will write a function to get all the player data from each table on the Wikipedia page:

```{r}
# function to get table for each team
get_teams <- function(tbl_nums) {
  
  # get country names
  team_name <-   url %>%
    read_html() %>%
    html_node(xpath = paste0('//*[@id="mw-content-text"]/div/h3[',tbl_nums,']')) %>%
    html_text
  
  # get squad info
  url %>%
    read_html() %>%
    html_node(xpath = paste0('//*[@id="mw-content-text"]/div/table[',tbl_nums,']')) %>%
    html_table(fill = TRUE) %>%
    as.tibble() %>%
    mutate(team_name = team_name) # add country names
}
```

Then, we will use `map` to iterate over our vector of numbers and pull player data for each team:

```{r}
teams <- map_df(table_numbers, get_teams)
```

We will be stacking all the data for 2018 squads on top of all the historical team data for analysis later. As a result we need to make some changes to our extracted data so we can `bind_rows` with the historical data.

```{r}
# clean up the data
teams <- teams %>%
  filter(!is.na(Caps)) %>%
  select(-`0#0`) %>%
  # add and change columns to match historical data
  mutate(No = NA, ClubCountry = NA, Year = 2018) %>%
  rename(Pos = `Pos.`, `DOB/Age` = `Date of birth (age)`, Country = team_name) %>%
  select(-Goals) %>%
  select(No,Pos,Player,`DOB/Age`,Caps,Club,Country,ClubCountry,Year)
```