---
title: World Cup Player Caps and Clubs
author: Michael Pawlus
date: '2018-06-12'
slug: world-cup-player-caps-and-clubs
categories:
  - rstats
tags:
  - Soccer
---

# World Cup Player Caps and Clubs

This is post three in a series using data for all World Cup teams collected in the [first post](https://michaelpawlus.netlify.com/2018/06/05/all-world-cup-squads/)

In the [second post](https://michaelpawlus.netlify.com/2018/06/08/age-of-world-cup-players/) we looked explored the ages of players over time, by team and by position as well as checked to see if there was a skewed distribution among birth months.

For this post we will look at caps which is the term used for counting how many appearances a player has made for games at the international level. We will also look at which clubs have the most players competing at this World Cup.

### Setup
```{r message=FALSE, warning=FALSE}
library(tidyverse) # for data manipulation
library(plotly) # for interactive plots
library(knitr) # for tables

# load data
all_squads <- read_csv("https://raw.githubusercontent.com/michaelpawlus/worldcup2018/squads/squads.csv")
```

### Caps Over Time

Here we will look at how many caps players have accumulated by the time they participate in a World Cup:

```{r message=FALSE, warning=FALSE}
## trend in caps over time
ggplot(all_squads, aes(x=Year, y=Caps)) +
  geom_jitter(alpha=0.2, color="#171714") +
  stat_summary(fun.y=mean, geom="line", size=1, color="#D30208") +
  stat_summary(fun.y=median, geom="line", size=1, color="#E5C685") +
  # note on mean line
  annotate("rect", xmin = 1928, xmax = 1932, ymin = 130, ymax = 140,
           color = "#D30208", fill = "#D30208") +
  annotate("text", x = 1934, y = 135, label = "average caps", color="#171714", hjust = 0, size = 4) +
  # note on median line
  annotate("rect", xmin = 1928, xmax = 1932, ymin = 110, ymax = 120,
           color = "#E5C685", fill = "#E5C685") +
  annotate("text", x = 1934, y = 115, label = "median caps", color="#171714", hjust = 0, size = 4) +
  theme(panel.background = element_blank(), 
        panel.grid.major.y = element_line(color = "#171714", size = 0.25), 
        legend.position="none",
        plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Year", y = "Caps", title = "World Cup Player Caps over time")
```

This number has been trending up. Are there just more competitions now then there used to be? Plus, we already saw that older and older players are competing. It is somewhat arbitrary but I am interested in seeing who those dots are above the 150 line for 2018.

### Those with 150 or more Caps playing this year

```{r}
all_squads %>%
  filter(Year == 2018, Caps > 149) %>%
  select(Year,Player,Pos,Country,Club,Caps) %>%
  arrange(desc(Year),-Caps) %>%
  distinct(Player, .keep_all = TRUE) %>%
  kable()
```

### Caps by Team

Next, we can look at caps by team like we did with age by team.

```{r message=FALSE, warning=FALSE}
wc18_adj_age <- all_squads %>%
  filter(Year == 2018) %>%
  mutate(Country = fct_reorder(Country, desc(Country)), Pos = fct_relevel(str_extract(Pos,"[:alpha:]+"), "GK","DF","MF","FW"))

wc18_adj_cap_plot <- ggplot(wc18_adj_age, aes(Country, Caps)) +
  geom_point(color = "#015386") + 
  stat_summary(fun.y = min, colour = "#D30208", geom = "point", size = 3, shape=21, show.legend = TRUE) +
  stat_summary(fun.y = max, colour = "#E5C685", geom = "point", size = 3, shape=21, show.legend = TRUE) +
  stat_summary(fun.y = mean, colour = "#D30208", geom = "point", size = 3, shape=23, show.legend = TRUE) +
  stat_summary(fun.y = median, colour = "#E5C685", geom = "point", size = 3, shape=23, show.legend = TRUE) +
  coord_flip() +
  theme(text = element_text(color = "#171714"),
        panel.background = element_blank(), 
        panel.grid.major.x = element_line(color = "#171714", size = 0.15), 
        legend.position="none",
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Caps for World Cup 2018 Teams")

wc18_adj_cap_plot <- ggplotly(wc18_adj_cap_plot)

text_vec <- paste0("Country: ",wc18_adj_age$Country,"<br />Name: ",wc18_adj_age$Player,"<br />Position: ",wc18_adj_age$Pos,"<br />Caps: ",wc18_adj_age$Caps)

wc18_adj_cap_plot$x$data[1][[1]][[3]] <- text_vec

wc18_adj_cap_plot
```

We can see that England, Morocco and Senegal all have no players with more than 60 caps.

Mexico also stand out in that the player with the fewest caps on their team is Edson Ãlvarez with 13. Every other team has at least one player with under 5 caps.

All CONCACAF teams (Costa Rica, Mexico, Panama) have players with over 50 caps on average. Does this highlight the lack of depth for countries from thos conference?

### Player Count by Club

Lastly, we can look at a count of players by their clubs. 

```{r}
# count clubs
all_squads %>%
  filter(Year == 2018) %>%
  group_by(Club) %>%
  tally() %>%
  arrange(n) %>%
  top_n(11) %>%
  ggplot(aes(x = factor(Club,levels=Club,ordered=TRUE), y = n)) +
  geom_point(color = "#171714", size = 5) +
  scale_y_continuous(breaks=seq(8,17,1)) +
  coord_flip() +
  theme(text = element_text(color = "#171714"),
        panel.background = element_blank(), 
        panel.grid.major.x = element_line(color = "#171714", size = 0.15), 
        legend.position="none",
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Clubs", y = "Players in WC Squads", title = "World Cup Player Count by Club (Top 12)")
```

Manchester City have sent the largest delegation. However, we see four of the top six teams from the Premier League sending 10+ players. Absent from the list are Arsenal (who are rebuilding) and Liverpool. What does this portend for Liverpool next season? 