---
title: World Cup Player Caps and Clubs
author: Michael Pawlus
date: '2018-06-12'
slug: world-cup-player-caps-and-clubs
categories:
  - rstats
tags:
  - Soccer
---

# World Cup Player Caps and Clubs

This is post three in a series using data for all World Cup teams collected in the [first post](https://michaelpawlus.netlify.com/2018/06/05/all-world-cup-squads/)

In the [second post](https://michaelpawlus.netlify.com/2018/06/08/age-of-world-cup-players/) we looked explored the ages of players over time, by team and by position as well as checked to see if there was a skewed distribution among birth months.

For this post we will look at caps which is the term used for counting how many appearnces a player has made for games at the international level. We will also look at which clubs have the most players competing at this World Cup.

### Setup
```{r}
library(tidyverse) # for data manipulation
library(plotly) # for interactive plots
library(knitr) # for tables

# load data
all_squads <- read_csv("https://raw.githubusercontent.com/michaelpawlus/worldcup2018/squads/squads.csv")
```

### Caps Over Time

```{r}
## trend in caps over time
ggplot(all_squads, aes(x=Year, y=Caps)) +
  geom_jitter(alpha=0.2, color="#171714") +
  stat_summary(fun.y=mean, geom="line", size=1, color="#D30208") +
  stat_summary(fun.y=median, geom="line", size=1, color="#E5C685") +
  # note on mean line
  annotate("rect", xmin = 1928, xmax = 1932, ymin = 130, ymax = 140,
           color = "#D30208", fill = "#D30208") +
  annotate("text", x = 1934, y = 135, label = "average caps", color="#171714", hjust = 0, size = 4) +
  # note on median line
  annotate("rect", xmin = 1928, xmax = 1932, ymin = 110, ymax = 120,
           color = "#E5C685", fill = "#E5C685") +
  annotate("text", x = 1934, y = 115, label = "median caps", color="#171714", hjust = 0, size = 4) +
  theme(panel.background = element_blank(), 
        panel.grid.major.y = element_line(color = "#171714", size = 0.25), 
        legend.position="none",
        plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Year", y = "Caps", title = "World Cup Player Caps over time")
```

### Those with 150 or more Caps playing this year

```{r}
all_squads %>%
  filter(Year == 2018, Caps > 149) %>%
  select(Year,Player,Pos,Country,Club,Caps) %>%
  arrange(desc(Year),-Caps) %>%
  distinct(Player, .keep_all = TRUE) %>%
  kable()
```

### Caps by Team

```{r}
wc18_adj_age <- all_squads %>%
  filter(Year == 2018) %>%
  mutate(Country = fct_reorder(Country, desc(Country)), Pos = fct_relevel(str_extract(Pos,"[:alpha:]+"), "GK","DF","MF","FW"))

wc18_adj_cap_plot <- ggplot(wc18_adj_age, aes(Country, Caps)) +
  geom_point(color = "#015386") + 
  stat_summary(fun.y = min, colour = "#D30208", geom = "point", size = 3, shape=21, show.legend = TRUE) +
  stat_summary(fun.y = max, colour = "#E5C685", geom = "point", size = 3, shape=21, show.legend = TRUE) +
  stat_summary(fun.y = mean, colour = "#D30208", geom = "point", size = 3, shape=23, show.legend = TRUE) +
  stat_summary(fun.y = median, colour = "#E5C685", geom = "point", size = 3, shape=23, show.legend = TRUE) +
  coord_flip() +
  theme(text = element_text(color = "#171714"),
        panel.background = element_blank(), 
        panel.grid.major.x = element_line(color = "#171714", size = 0.15), 
        legend.position="none",
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Caps for World Cup 2018 Teams")

wc18_adj_cap_plot <- ggplotly(wc18_adj_cap_plot)

text_vec <- paste0("Country: ",wc18_adj_age$Country,"<br />Name: ",wc18_adj_age$Player,"<br />Position: ",wc18_adj_age$Pos,"<br />Caps: ",wc18_adj_age$Caps)

wc18_adj_cap_plot$x$data[1][[1]][[3]] <- text_vec

wc18_adj_cap_plot
```

### Player Count by Club

```{r}
# count clubs
all_squads %>%
  filter(Year == 2018) %>%
  group_by(Club) %>%
  tally() %>%
  arrange(n) %>%
  top_n(11) %>%
  ggplot(aes(x = factor(Club,levels=Club,ordered=TRUE), y = n)) +
  geom_point(color = "#171714", size = 5) +
  scale_y_continuous(breaks=seq(8,17,1)) +
  coord_flip() +
  theme(text = element_text(color = "#171714"),
        panel.background = element_blank(), 
        panel.grid.major.x = element_line(color = "#171714", size = 0.15), 
        legend.position="none",
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Clubs", y = "Players in WC Squads", title = "World Cup Player Count by Club (Top 12)")
```