---
title: World Cup 2018 Predicted Results
author: Michael Pawlus
date: '2018-06-14'
slug: world-cup-2018-predicted-results
categories:
  - rstats
tags:
  - Soccer
---

# World Cup 2018 Predicted Results

This is post four in a series using data for all World Cup teams collected in the [first post](https://michaelpawlus.netlify.com/2018/06/05/all-world-cup-squads/)

In the [second post](https://michaelpawlus.netlify.com/2018/06/08/age-of-world-cup-players/) we looked explored the ages of players over time, by team and by position as well as checked to see if there was a skewed distribution among birth months.

In [post three](https://michaelpawlus.netlify.com/2018/06/12/world-cup-player-caps-and-clubs/) we looked at the number of caps for players over time and by team as well as the clubs sending the largest delegations.

This post is just to get the results published. I will add another post soon with how we arrived at these results.

I will just say that we did not use a very complex approach due to time constraints and mostly predicted winners and losers based on a couple comparisons of team strength. The goal difference was not really calculated but are just approximated here due to some teams being level on points.

```{r message=FALSE, warning=FALSE, echo=FALSE}
# load libraries
library(tidyverse) # for data manipulation
library(knitr) # for tables

# load data
group_tables <- read_csv("https://raw.githubusercontent.com/michaelpawlus/worldcup2018/squads/group_stage_table.csv")
round_of_16 <- read_csv("https://raw.githubusercontent.com/michaelpawlus/worldcup2018/squads/knockout_round16_matches.csv")
quarters <- read_csv("https://raw.githubusercontent.com/michaelpawlus/worldcup2018/squads/knockout_semi_matches.csv")
semis <- read_csv("https://raw.githubusercontent.com/michaelpawlus/worldcup2018/squads/knockout_semi2_matches.csv")
final <- read_csv("https://raw.githubusercontent.com/michaelpawlus/worldcup2018/squads/knockout_final_match.csv")
```

## Group Stage Predictions

### Group A:
```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group A") %>%
  kable()
```

### Group B:
```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group B") %>%
  kable()
```

### Group C:
```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group C") %>%
  kable()
```

### Group D:
```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group D") %>%
  kable()
```

### Group E:

Note: In this group we have Switzerland finishing ahead of Serbia.

```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group E") %>%
  kable()
```

### Group F:
```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group F") %>%
  kable()
```

### Group G:
```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group G") %>%
  kable()
```

### Group H:
```{r message=FALSE, warning=FALSE, echo=FALSE}
group_tables %>%
  filter(group == "Group H") %>%
  kable()
```

## Knockout Rounds

### Round of 16
```{r message=FALSE, warning=FALSE, echo=FALSE}
round_of_16 %>%
  select(o_team,d_team,team_a) %>%
  rename(`Team A` = o_team, `Team B` = d_team, Winner = team_a) %>%
  kable()
```

### Quarter-Finals
```{r message=FALSE, warning=FALSE, echo=FALSE}
quarters %>%
  select(o_team,d_team,team_a) %>%
  rename(`Team A` = o_team, `Team B` = d_team, Winner = team_a) %>%
  kable()
```

### Semi-Finals
```{r message=FALSE, warning=FALSE, echo=FALSE}
semis %>%
  select(o_team,d_team,team_a) %>%
  rename(`Team A` = o_team, `Team B` = d_team, Winner = team_a) %>%
  kable()
```

### Final
```{r message=FALSE, warning=FALSE, echo=FALSE}
final %>%
  select(o_team,d_team,team_a) %>%
  rename(`Team A` = o_team, `Team B` = d_team, Winner = team_a) %>%
  kable()
```
